Wigglyhedra

### 2024-06-12 draft/wigglyhedra.tex
minor changes

\subsection{A triangulated category associated to a quiver}
We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra~\(A_m\), called the \defn{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i|j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1|i_2|\cdots|i_n)\).

\begin{definition}
  The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the following relations:
  \begin{itemize}
  \item \((i|j|l) = 0\) whenever \(i \neq l\);
  \item \((i|j|i) = (i|l|i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
  \vincent{why not write $i-1$ and~$i+1$?}
  \end{itemize}
\end{definition}

@@

@@

We refer the reader to e.g.~\cite[1.3]{bel.ber.del:82} for the definition of a \(t\)-structure on a triangulated category.
The following proposition is easily checked using the calculations in~\cref{rem:generating-morphisms} and standard techniques involving categories of complexes; we omit the proof.
\begin{proposition}
  The pair of subcategories \((\mathcal{C}_m^{\leq 0}, \mathcal{C}_m^{\geq 0})\) forms a bounded \(t\)-structure on \(\mathcal{C}_m\).
\end{proposition}
We call this \(t\)-structure the \defn{standard \(t\)-structure} on \(\mathcal{C}_m\).
Set \(\mathcal{A}_m = \mathcal{C}^{\leq 0}_m \cap \mathcal{C}_m^{\geq 0}\); this is the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, which contains all linear complexes.
It follows that \(\mathcal{A}_m\) is abelian.

Also record the following statements for future use; the first may be found in~\cite[1.3]{bel.ber.del:82}
\vincent{is that a section?}
\begin{defprop}\label{defprop:cohomology}
  Let \((\mathcal{C}^{\le 0}, \mathcal{C}^{\geq 0})\) be a \(t\)-structure on a triangulated category \(\mathcal{C}\) with heart~\(\mathcal{A} = \mathcal{C}^{\leq 0} \cap \mathcal{C}^{\geq 0}\).
  For each integer \(i\), we have the \(i\)th truncation functors
  \[\tau^{\leq i} \colon \mathcal{C} \to \mathcal{C}^{\leq i} \quad\text{and}\quad \tau^{\geq i} \colon \mathcal{C} \to \mathcal{C}^{\geq i},\]
  which are right and left adjoint respectively to the inclusion functors \(\mathcal{C}^{\leq 0} \to \mathcal{C}\) and \(\mathcal{C}^{\geq 0} \to \mathcal{C}\).
  The \(i\)th \defn{cohomology functor} is defined to be the composition \(\tau^{\leq i} \circ \tau^{\geq i}\).
  It is canonically isomorphic to the composition \(\tau^{\geq i} \circ \tau^{\leq i}\), and maps an object of \(\mathcal{C}\) to an object of \(\mathcal{A}\).
\end{defprop}
The next statement, easy to check from the definitions, is mentioned in~\cite[Sect.~3.3]{bayer2011tour}.
\begin{defprop}\label{defprop:cohomology-filtration}
  Let \(\mathcal{C}\) be any triangulated category, and \(\mathcal{A} \subset \mathcal{C}\) the heart of a bounded \(t\)-structure on \(\mathcal{C}\).
  Let \(X\) be a non-zero object of \(\mathcal{C}\), and let \(k\) and \(l\) be the maximum and minimum integers respectively for which \(H^k(X)\) and \(H^{l}(X)\) are non-zero.
  Then \(X\) has a unique \defn{cohomology filtration}, which is the unique filtration of the form
  \[
    \begin{tikzcd}[column sep=0.1cm]
      0  \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k+1} \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]  
\end{defprop}

@@

\subsection{Categorical constraints on compatibility of curves}
In this section, we give a categorical criterion for two wiggly arcs to be compatible as in~\cref{def:compatible}.
As such, this section is somewhat dual to the previous section --- the criterion in this section determines whether a pair of wiggly arcs can belong together in a single wiggly pseudotriangulation, whereas the previous section generated a wiggly pseudotriangulation given any object of the category.
\begin{proposition}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X\langle g \rangle, X\langle g' \rangle) = 0\) for any two integers \(g\) and \(g'\).
\end{proposition}
\begin{proof}
  \asilata{TODO.}
\end{proof}
\begin{remark}
  This proposition is similar to the condition of being compatible in a cluster-tilting object.
  \asilata{Expand and make precise.}
\end{remark}
\begin{remark}
  As stated, this proposition is not expected to be true in greater generality.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
  \asilata{Decide what to say here, because the forthcoming work is far from complete, although I do have counterexamples in type \(D_4\). Discussing how to interpret these counterexamples may not involve too much more work?}
\end{remark}
\begin{question}
  How to best relate this section to the previous one purely categorically?
  That is, given the categorical constraint on compatibility of arcs, is it also clear that it gives a constraint on objects appearing in the cohomology filtration?
  \asilata{Make more precise.}
\end{question}

>>>>

\subsection{A triangulated category associated to a quiver}
We begin by describing the category of interest, which is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra~\(A_m\), called the \defn{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i, j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1, i_2, \dots, i_n)\).

\begin{definition}
  The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the relations: \((i-1, i, i+1) = 0 = (i+1, i, i-1)\) and \((i, i-1, i) = (i, i+1, i)\) for all~\(1 < i < m\).
%  \begin{itemize}
%  \item \((i-1, i, i+1) = (i+1, i, i-1\) for all~\(1 < i < m\),
%  \item \((i, i-1, i) = (i, i+1, i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
%  \end{itemize}
\end{definition}

@@

\vincent{What about this remark instead?}
\begin{remark}\label{rem:generating-morphisms}
  It is easy to check that the generating objects \(P_i\) (and hence their graded and homological shifts) satisfy \(\Hom^{0,0}(P_i,P_i) = \Hom^{1,0}(P_i,P_i) = \Hom^{1,0}(P_i,P_{i\pm1}) = k\) and \(\Hom^{g,h}(P_i, P_j) = 0\) in all other cases.
\end{remark}

@@

We refer the reader to e.g.~\cite[Sect.~1.3]{bel.ber.del:82} for the definition of a \(t\)-structure on a triangulated category.
The following proposition is easily checked using the calculations in~\cref{rem:generating-morphisms} and standard techniques involving categories of complexes; we omit the proof.
\begin{proposition}
  The pair of subcategories \((\mathcal{C}_m^{\leq 0}, \mathcal{C}_m^{\geq 0})\) forms a bounded \(t\)-structure on \(\mathcal{C}_m\).
\end{proposition}
We call this \(t\)-structure the \defn{standard \(t\)-structure} on \(\mathcal{C}_m\).
Set \(\mathcal{A}_m = \mathcal{C}^{\leq 0}_m \cap \mathcal{C}_m^{\geq 0}\); this is the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, which contains all linear complexes.
It follows that \(\mathcal{A}_m\) is abelian.

Also record the following statements for future use; the first may be found in~\cite[Sect.~1.3]{bel.ber.del:82}
\begin{defprop}\label{defprop:cohomology}
  Let \((\mathcal{C}^{\le 0}, \mathcal{C}^{\geq 0})\) be a \(t\)-structure on a triangulated category \(\mathcal{C}\) with heart~\(\mathcal{A} = \mathcal{C}^{\leq 0} \cap \mathcal{C}^{\geq 0}\).
  For each integer \(i\), we have the \(i\)th truncation functors
  \[\tau^{\leq i} \colon \mathcal{C} \to \mathcal{C}^{\leq i} \quad\text{and}\quad \tau^{\geq i} \colon \mathcal{C} \to \mathcal{C}^{\geq i},\]
  which are right and left adjoint respectively to the inclusion functors \(\mathcal{C}^{\leq 0} \to \mathcal{C}\) and \(\mathcal{C}^{\geq 0} \to \mathcal{C}\).
  The \(i\)th \defn{cohomology functor} is defined to be the composition \(\tau^{\leq i} \circ \tau^{\geq i}\).
  It is canonically isomorphic to the composition \(\tau^{\geq i} \circ \tau^{\leq i}\), and maps an object of \(\mathcal{C}\) to an object of \(\mathcal{A}\).
\end{defprop}
The next statement, easy to check from the definitions, is mentioned in~\cite[Sect.~3.3]{bayer2011tour}.
\begin{defprop}\label{defprop:cohomology-filtration}
  Let \(\mathcal{C}\) be any triangulated category, and \(\mathcal{A} \subset \mathcal{C}\) the heart of a bounded \(t\)-structure on \(\mathcal{C}\).
  Let \(X\) be a non-zero object of \(\mathcal{C}\), and let \(k\) and \(l\) be the maximum and minimum integers respectively for which \(H^k(X)\) and \(H^{l}(X)\) are non-zero.
  Then \(X\) has a unique \defn{cohomology filtration}, which is the unique filtration of the form
  \[
    \begin{tikzcd}[column sep=0.1cm]
      0  \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k+1} \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]  
\end{defprop}

@@

\subsection{Categorical constraints on compatibility of curves}
In this section, we give a categorical criterion for two wiggly arcs to be compatible as in~\cref{def:compatible}.
As such, this section is somewhat dual to the previous section --- the criterion in this section determines whether a pair of wiggly arcs can belong together in a single wiggly pseudotriangulation, whereas the previous section generated a wiggly pseudotriangulation given any object of the category.
\begin{proposition}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X\langle g \rangle, X\langle g' \rangle) = 0\) for any two integers \(g\) and \(g'\).
\end{proposition}
\begin{proof}
  \asilata{TODO.}
\end{proof}
\begin{remark}
  This proposition is similar to the condition of being compatible in a cluster-tilting object.
  \asilata{Expand and make precise.}
\end{remark}
\begin{remark}
  As stated, this proposition is not expected to be true in greater generality.
  For instance, it does not hold for the zigzag algebra of Dynkin type~$D_4$.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
  \asilata{Decide what to say here, because the forthcoming work is far from complete, although I do have counterexamples in type \(D_4\). Discussing how to interpret these counterexamples may not involve too much more work?}
\end{remark}
\begin{question}
  How to best relate this section to the previous one purely categorically?
  That is, given the categorical constraint on compatibility of arcs, is it also clear that it gives a constraint on objects appearing in the cohomology filtration?
  \asilata{Make more precise.}
\end{question}
